\documentclass[aagt.tex]{subfiles}
\begin{document}
\lecture{16.06.2016}

\begin{defi*}
  \emph{Upper neighbourhood} 
  \[ U\Gamma(v) \coloneqq \{ w \in \Gamma(v): \sigma(w) > \sigma(v) \} \]
  where $\sigma: V(G) \to \{1,2,\dots,n\}$ (bijective or ordering), $n \coloneqq \abs{V(G)}$.
\end{defi*}

\begin{prop}[6.9 Fulkerson, Gross 1965]
  $G=(V,E)$ is chordal if and only if there exists a $\sigma: V(G) \to \{1,2,\dots,n\}$ bijective, which is PES.
\end{prop}

\begin{proof}
  \begin{enumerate}
    \item[$\Rightarrow$] trivial by respected application of Lemma \ref{6.8-lemma}.
    There exists $v_1$ simplical; $G^\prime = G \setminus \{v_1\} \subseteq G$ is chordal, There exists a simplicial vertex $v_2$, $G^\prime = G \setminus \{v_1,v_2\} \subseteq G$ chordal, There exists a s. v. $v_2$ and so on.
    Claim: $v_1,v_2,\dots,v_n$ is PES.
    Indeed because $v_i$ is s. v. in $G \setminus \{v_1,\dots,v_{i-1} \}$, $1 \leq i \leq n$ and therefore $\Gamma_{G \setminus \{v_1,\dots,v_{i-1} \}} (v_i) = U\Gamma_G(v_i)$ is a clique.
    \item[$\Leftarrow$] Assume $G$ has a PES. Show $G$ is chordal.
    Let $C$ be a cycle with $\geq 4$ vertices.
    Let $u = arg\min \{ \inv{\sigma}(v): v \in V(C) \}$
    \todo{pic}{add pic}
    $\implies u^-, u^+ \in U\Gamma(u) \implies \{u^-,u^+\} \in E(G)$ which is a clique
  \end{enumerate}
\end{proof}

Algorithmic idea for recognition of chordal graphs

Set $i=0$ check whether  $G$ a simplical vertex. If not, stop \enquote{$G$ is not chordal}.
If yes, let $v_i$ be a s.v. Set $G = G \setminus \{v_i\}$, $i = i+1$.
Repeat until $i = n$.

Possible implementation in $\bigO(n^2 m)$ because for-loop is repeated $\leq n$ times.

Each repetition in $\bigO(nm)$: $G \coloneqq G \setminus \{v_i\}$ in $\bigO(\deg(v_i))$,
check for s.v. in $G$; there are at most $n$ candidates for a s.v.
The simplicity could be checked in $\bigO(m)$ times (need to check whether $\Gamma_G(v)$ contains all edges in $\bigO(m)$ time.

Speed up: 

R.E. Tarjan and M. Vannekakis, Simple linear time algorithms to test chordalty of graphs, but acyclicity of hypergraphs and selectivly reduce ancyclic hypergraphs, SIAM J. on Computing 13, 1984, 566-579;
Addendum 14, 1985, 254-255

\subsubsection*{2 step procedure}

\begin{enumerate}
  \item Construction of total ordering $\sigma$ in $V(G)$ such that $G$ is PES $\iff G$ chordal
  \item check whether $\sigma$ is PES.
\end{enumerate}

Ad 1) This \enquote{special} $\sigma$ is the reverse \emph{maximum adjecency order}

\begin{defi*}
  MA Order: $\sigma: V(G) \to \{1,2,\dots,n\}$ bijective such that for all $1 \leq i \leq n$:
  \[ \abs{ \{ v \in \{v_1,\dots,v_{i-1}\}: \{v_i,v\} \in E(G) \}} = \max_{u \notin \{v_1,\dots,v_{i-1}\}} \abs{ \{v \in \{v_1,\dots,v_{i-1}\} : \{u,v\} \in E(G)\}} \] \todo{text missing}
\end{defi*}

\begin{ex}
  \todo{pic}{add pic}
\end{ex}

Result from CombOpt 1 (or Korte and Vygens, Conbinatonal Optimisation Theory and Applications, 2010, Springer)

MA can be cmputed $\bigO(m+ n \log n)$ time by Fibonacci heaps.

\enquote{Special} $\sigma$ is the reversed MA order.

\begin{prop}[6.10] \label{6.10-prop}
  A total ordering $\sigma$ in $V(G)$ is a PES $\iff$ 
  If $v_i,v_j \in V(G)$ are connected by a path $P$ with inner vertices $n$ with $\sigma(n) < \min \{\sigma(v_i),\sigma(v_j)\}$, then $\{v_i,v_j\} \in E(G)$.
\end{prop}

\begin{proof}
  \begin{enumerate}
    \item[$\Leftarrow$] trivial: Consider $U\Gamma(v)$
    \todo{pic}{add pic}
    Path $P$ fulfills the property mentioned in the implication $\implies \{u_1,u_2\} \in E(G)$
    \item[$\Rightarrow$] Assume $\sigma$ is PES. Show the implication.
    Let $v_i,v_j \in V(G)$ be connected by path $P$ as in the implication. Want to show $\{v_i,v_j\} \in E(G)$.
    Assume wlog $P$ is the shortest such path.
    If $\{v_i,v_j\} \in E(G)$ $\checkmark$
    If not, there exists an inner vertex in $P$. Choose the inner vertex $u$ of $P$ with smallest $\sigma$-label.
    \todo{pic}{add pic}
    $u^-,u^+ \in U\Gamma(v) \overset{\sigma \text{ is PES}}{implies} \{u^-,u^+\} \in E(G) \implies P$ could be shortened $\lightning$ 
  \end{enumerate}
\end{proof}

\begin{prop}[6.11]
  Let $G$ be a chordal graph. Let MA order yield $v_1,\dots,v_n$.
  Then the vertices of $G$ in reversed order,
  $v_n,v_{n-1},\dots,v_1$ build a PES.
\end{prop}

\begin{proof}
  $\sigma: V(G)  \to \{1,2,\dots,n\}$ bijective. $\sigma(v_i) = n-i+1$
  
  We claim that property \ref{0616_star} holds:
  \begin{align}\label{0616_star}
    \text{If } \sigma(u) < \sigma(r) < \sigma(w) \text{ and } \{u,w \} \in E(G), \{v,w\} \notin E(G) \text{, then } \exists z \in V(G), \\
    \text{such that } \sigma(v) < \sigma(z), \{v,z\} \in E(G), \{u,u\} \notin E(G)
  \end{align}
  Indeed: When MA labels $v$ (which is not connected to $w$ while $u$ is), there must be a labelled vertex which is a neighbour of $v$ but not a neighbour of $u$; call it $z$ and check the properties in \ref{0616_star}.
  
  Assume $\sigma$ is not a PES. By Propositon \ref{6.10-prop}, there exist $v_i,v_j \notin E(G)$ such that there exists a $v_i$-$v_j$-path $P$ with 
  \begin{align} \label{0616_property}
    \text{inner vertices } u \text{ with } \sigma(u) < \min \{\sigma(v_i),\sigma(v_j)\} \text{.}
  \end{align}
  Select such a pair $v_i,v_j$ where $v_j$ has the largest label $\sigma(v_j)$.
  If there are many possible vertices $v_i$ select the one with largest $\sigma$.
  Let $P$ be a shortest $v_i$-$v_j$-path with property \ref{0616_property}.
  \todo{pic}{add pic}
  Since $\{v_i,v_j\} \in E(G) \implies \exists$ inner vertex of $P$, let $u$ be inner vertex adjacent to $v_j$.
  Apply property \ref{0616_star} to $u,v_i,v_j$ with $\sigma(u) < \sigma(v_i) < \sigma(v_j)$.
  $\implies \exists z: \sigma(z) > \sigma(v_i)$, $\{z,v_i\} \in E(G), \{u,z\}\notin E(G)$.
  $\{z,v_i\} \in E(G)$; otherwise path $z$-$v_j$ would fulfill condition of \ref{0616_star} and contradict the selection of $v_i$.
  
  $P \cup \{z,v_j\}$ is a cycle $C$ with $\geq 4$ vertices $\overset{G \text{ chordal}}{\implies} \exists$ chordal in $C$, chord has to pass through $z$ otherwise $P$ was not shortest.
  
  Let $x$ be the first vertex in $P$ (starting at $v_j$) which is connected $z$, $x$ has to the left of $n$ ($\{z,u\} \notin E(G)$)
  But then there is new cycle $C^\prime$ (length $\geq 4$) without a chord $\lightning$.
\end{proof}

\begin{theorem}
  It can be checkd in $\bigO(n + m)$ time if the reversed MA order is a PES.
\end{theorem}

\begin{proof}
  No proof.
\end{proof}

\begin{cor}
  This leads to an $\bigO(m + n \log n)$ algorithm for recognition of chordal graphs.
\end{cor}

\begin{rem}
  The number $\alpha(G)$, $\omega(G)$, $\chi(G)$, $\theta(G)$ can be computed in $\bigO(m + n \log n)$ time in chordal graphs (and in efficient time n perfect graphs).
\end{rem}

Consider a PES $\sigma$ (reversed MA order)
Colouring: Greedy with order $\sigma$ leads to an optimal colouring (exercises).
Clique: All max cliques (wrt inclusion) are among $\{v\} \cup U\Gamma(v)$ (exercises).
Stability: Set $s_1 = \inv{\sigma}(1)$; inductively set $s_k$ to be the first vertex after $s_{k-1}$ which has no neighbours among $s_1,\dots,s_{k-1}$.
Consider $S \coloneqq \{s_1,\dots,s_\ell\}$, it is a stable set and also a max stable set because $(\{s_k\} \cup U\Gamma(s_k))_{1 \leq k \leq \ell}$ is a family of cliques such that
\[ V(G) = \bigcup_{k=1}^\ell  \{\{s_k\} \cup \Gamma(s_k)\} \text{.} \]

So we found a stable set $S$ with $\abs{S} = \ell$ and a covering by cliques  with $\ell$ cliques.
$\max \abs{S} \leq \min$ number of cliques in clique covering

$\max_{S \text{ stable}} \abs{S} \leq $ min cardinality of a clique covering, i.e. 
\[ \alpha(G) \leq \theta(G) \]
Implies $S=\{s_1,\dots,s_\ell\}$ and $(\{s_k\} \cup U \Gamma(s_k))_{1 \leq k \leq \ell}$ is a certificate of optimality.


\end{document}
